\thetitle

### 2021-03-01 writing/mutate-benchmarks.tex
adding prose for mutators design and benchmarks selection

Over the past decade, \citet{gtnffvf-jfp-2019} have curated a collection of
Typed Racket programs for systematically measuring the implementation's
performance. The benchmark suite consists of fully typed, correct programs,
written by a number of different authors, who have maintained and evolved these
programs over several years. The programs range widely in size, complexity,
purpose, and in programming style. The latter is mostly a reference to their
reliance on many Typed Racket features: occurrence typing~\cite{tf-icfp-2010},
types for mutable and immutable data structures~\cite{hpst-sfp-2010}, types for
first-class classes and objects~\cite{tsdtf-oopsla-2012}, and types for Racket's
numeric tower~\cite{stathff-padl-12}.

Since mixing of typed and untyped code in Typed Racket takes place at the module
level, the experiment calls for benchmarks with a decent number of modules and a
variety of module dependency graphs. Filtering the benchmark suite using these
criteria yields ten suitable programs. Figure~\ref{table:benchmark-descriptions}
displays the selected benchmark programs and describes their basic
characteristic.

>>>>

Since this is the first empirical attempt to evaluate blame in gradual
typing, there is no standard benchmark suite that we can use off-the-self.
In addition, there is no exsitng benchmark suite of gradually typed
programs with type-level mistakes that we could
repurpose.\footnote{Recently ~\citet{cc-oopsla-20} created a
Reticulated Python program collection to evaluate their technique for
fixing mistakes in type annotations; unfortunately their collection does
not come with type-level mistakes in the code proper.} In fact, the bar
for testing the rational programmer is much higher than the existence of a
collection of programs with type-level errors. Testing the rational
programmer demands a currated collection of programs that take advantage
of the sophisticated features of gradually typed languages such as Typed
Racket.

Over the past decade,  \citet{gtnffvf-jfp-2019} have curated such a
collection of Typed Racket programs for systematically measuring the
implementation's performance. The benchmark suite consists of fully typed,
correct programs, written by a number of different authors, who have
maintained and evolved these programs over several years. The programs
range widely in size, complexity, purpose, origin and in programming
style. The latter is mostly a reference to their reliance on many Typed
Racket features: occurrence typing~\cite{tf-icfp-2010}, types for mutable
and immutable data structures~\cite{hpst-sfp-2010}, types for first-class
classes and objects~\cite{tsdtf-oopsla-2012}, and types for Racket's
numeric tower~\cite{stathff-padl-12}. Hence they offer a diverse lignustic
substrate suitable for injecting type-level errors and testing the
rational programmer.

Since mixing of typed and untyped code in Typed Racket takes place at the
module level, the experiment calls for benchmarks with a decent number of
modules and a variety of module dependency graphs. Filtering the benchmark
suite using these criteria while preserving its linguistic diversity
yields ten suitable programs. Figure~\ref{table:benchmark-descriptions}
displays the selected benchmark programs and describes their basic
characteristic.

### 2021-03-01 writing/mutate-mutators.tex
adding prose for mutators design and benchmarks selection

For the evaluation of a blame strategy, the mutator must produce type-level
mistakes that the run-time checks of gradual typing systems or the safety
checks of the underlying language can detect. Once detected, the rational
programmer should be able to detect the mistake by gradually adding types to
blamed modules.

Figure~\ref{table:mutation-ops} describes the 16 mutators used for the
experiment.  They are inspired by the authors decades-long experience
of making type-level mistakes in Typed Racket, some of which take a
non-trivial effort to debug.  For each mutaror, the figure provides a
short description and an example mutation.

Most mutators are self-explanatory.  The first twelve apply to most
gradually typed languages, including those with classes.  The last
four target distinguishing features of Typed Racket.  Specifically,
{\tt arithmetic} may replace a {\tt +} with a {\tt -} in an attempt to
change the type of the result of the arithmetic operation. In Typed
Racket, {\tt +}'s result is a {\tt Positive-Integer} when all
arguments are positive integers~\cite{stathff-padl-12}. However, the
result of {\tt -} is {\tt Integer}. In the same spirit, the {\tt
boolean} mutator aims to take advantage of Typed Racket faithfulness
to Racket's ``falsiness.'' Finally, {\tt negate-cond} and {\tt
force-cond} attempt to confuse occurrence typing.

>>>>

For the evaluation of a blame strategy, mutators must produce type-level
mistakes that the run-time checks of gradual typing systems or the safety
checks of the underlying language can detect. Once detected, the rational
programmer should be able to locate the mistake by gradually adding types to
blamed modules.

Figure~\ref{table:mutation-ops} describes the 16 mutators used for the
experiment.  For each mutator, the figure provides a short description and
an example mutation.  Most are self-explanatory.  The first twelve apply
to most gradually typed languages, including those with classes.  The last
four target distinguishing features of Typed Racket.  Specifically, {\tt
arithmetic} may replace a {\tt +} with a {\tt -} in an attempt to change
the type of the result of the arithmetic operation. In Typed Racket, {\tt
+}'s result is a {\tt Positive-Integer} when all arguments are positive
integers~\cite{stathff-padl-12}. However, the result of {\tt -} is {\tt
Integer}. In the same spirit, the {\tt boolean} mutator aims to take
advantage of Typed Racket faithfulness to Racket's ``falsiness.'' Finally,
{\tt negate-cond} and {\tt force-cond} attempt to confuse occurrence
typing.

For the design of these mutators, the starting point is the mutators of
~\citet{lksfd-popl-2020} which in turn are borrowed from   the vast
literature on mutation testing~\cite{jia2011analysis}. However, only 2
of the previous mutators transfer as is to the context of this study, 1 is
entirely ineffective and 5 need significant modifications. For the
latter  the authors relied on their decades-long experience of making type-level mistakes in Typed
Racket, some of which take a non-trivial effort to debug.  In more detail,
the relational mutator of ~\citet{lksfd-popl-2020} does not lead to
type-level errors that Typed Racket's type system can detect, so it has
been eliminated. Similarly, the original arithmetic mutator does not
reliably lead to type errors, so it has been adapted to only make operator
swaps that can affect the type of the result. For example, changing a `/`
to `*` does not affect the type of an expression, but changing `*` to `/`
may s division by zero is detectable by Typed Racket's type system. 
In addition to careful tweaking of previous mutators, the final
set of mutators comes with 9 new ones that collectively
produce a sufficiently large and diverse set of type-level mistakes. 
For example, unlike the traditional subexpression swap mutator, 
the {\tt -id} family of mutators apply in different
contexts and utilize different pieces of non-local information about the
identifiers in a program.

As a final remark, the 16 selected mutators not only produce  type-level
mistakes  reliably but also interesting debugging scenarios. Most
mutators,  including variants of the ones
~\citet{w:good-types-mutation} propose, some times achieve the first but 
rarely the second. 
 For example,  adding a new public method to
a class surprisingly leads to mutants that either terminate 
successfully or to mutants that fail 
 immediately during the evaluation of the mutated class.  
 In contrast, the seemingly similar mutator that turns a private method into    
 a public one affects inheritance and leads to subtle bugs.  
Hence, due to the elusive nature of mutators, it is imperative to define
precisely what makes a debugging scenario interesting.
